---
title: "How to characterize the PROMISE dataset"
author: "Luke Johnston"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to characterize the PROMISE dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Prevalent and incident cases

PROMISE is a prospective cohort of individuals who are followed over time to see
who developed type 2 diabetes. So part of describing the dataset is determining
prevalence and incidence of diabetes, as well as general dysglycemia disorders
(impaired fasting glucose and impaired glucose tolerance). There are other
disease/disorders to characterise as well, which the functions explained below
will be able to calculate. Before we begin, let's first load up the package as
well as the example dataset.

```{r setup}
library(PROMISE.desc)
data("example_data")
dplyr::glimpse(example_data)
```

The first function `desc_prevalence` calculates prevalent cases of any
disease/disorder. We use it by selecting the columns we want to determine
prevalent cases. *A note*, the dataset *must* contain the columns `SID` and
`VN`.

```{r}
prev_data <- example_data %>% 
    desc_prevalence(c("DM", "IFG")) 
prev_data
```

This function more or less counts disease/disorder status at each time point and
re-arranges it into a wide format. If you want to convert this into a table, use:

```{r}
prev_data %>% 
    desc_table(caption = "*Example* dataset showing prevalence of DM and IFG at each time point.")
```

For incident cases, you can either add the incidence cases directly to the
dataset or describe (count) incident cases like above:

```{r}
example_data %>% 
    add_incidence(c("DM", "IFG", "IGT"), prefix = "incid_")
```

This adds the incident data to the original dataset. If you want to describe
(count) the incident cases, use:

```{r}
incid_data <- example_data %>% 
    desc_incidence(c("DM", "IFG"), prefix = "incid_")
incid_data
```

This function does the same thing as the `desc_prevalence` function, except it
adds the incident cases before and counts these cases. Again, if you want to
show this as a table, use:

```{r}
incid_data %>% 
    desc_table(caption = "*Example* dataset showing incident cases of DM and IFG at each time.")
```

